Binning Strategies and Related Loss for Binned Scatterplots of Large Data

### 2016-12-03 AuthorResponseToReviewers.Rmd

*I also fail to understand how the discussion on artificial striping in Section 5.1 is relevant for spatial loss.*

- Artificial striping is a symptom of poorly aligned bin sizes and the data resolution, which will lead to imbalanced bin allocation and thus more spatial loss. A clarification of this idea and a visual comparison of bin sizes and spatial loss have been added to section 5.1 in the context of Figure 6.

*Please provide quantified results on spatial loss when you talk about how binning origin affects spatial loss in Figure 7.*

- The statement addressing the Figure already contains a quantified comparison of spatial loss under the two scenarios. **"As an example, Figure 7 displays the binned scatterplots of ten-by-ten unit bins for the fine exponential data where the binning origin at (0,0) incurs seven percent lower spatial loss than for the (-9,-9) origin due to the heavy overlap into negative regions"** The difference is given as a percent decrease in spatial loss because the relative spatial loss change seemed most pertinent.

@@

*In visualization practices, it is common to use both ordinal scales as well as continuous scales when mapping numbers to colors. How might you address the issue of frequency loss when continuous scales are used?*

- The need for future perception research to quantify the implicit frequency loss when using continuous shading scales is now noted in the discussion section of the paper.

@@

*I don’t buy the advice to “use between four and seven frequency”. There simply isn’t empirical evidence in the paper to support this claim.*

- We do not want to allow more color bin than we can *easily* differentiate because the our expressed purpose for working with discretized color scales in this scenario is to have made loss of frequency information *entirely* explicit. While working with discretizing shades we found ourselves worrying about our ability to perfectly match bins to a legend after approximately 6 shades. We do not claim to have gathered empirical evidence, rather we defer to Healey and Enn's Theory on discernable colors (typically 4-7) to back out experience. Figure 9 was included for the reader to assess if they felt our suggestion was justifiable.

### 2016-12-03 versionForJCGS/PaperJCGS.Rnw
Added note on future work necessary to quantify implicit frequency loss. Also added rebuttal to reviewer 2's scoff at advice to restrict number of frequence bins.
Clarify and soften rebuttal of 4-7 scoff slightly.

If we elect to use frequency binning to shade tiles ordinally, it is recommended to use between four and seven distinct shades. This is done to minimize the frequency loss within the bounds of human perceptual ability to distinguish multiple shades simultaneously. Using quantile frequency binning or standard log frequency binning are shown to be reasonable methods -- with slight differences of interpretability -- for handling situations with heavily skewed bin counts.

>>>>

\ktm{Due to imperfect human perception may elect to use frequency binning to discretize the shading scale with the goal to make the loss of frequency information more explicit. If we aim to have highly accurate perceptitual mapping,} it is recommended to use at most between four and seven distinct shades. This is done to minimize the frequency loss within the bounds of human perceptual ability to distinguish multiple shades simultaneously. Using quantile frequency binning or standard log frequency binning are shown to be reasonable methods -- with slight differences of interpretability -- for handling situations with heavily skewed bin counts. \ktm{Frequency binning to discretize the shading scale allows us to explicitly quantify the loss of frequency information, however future perception research is needed to quantify the implicit frequency loss in reading binned scatterplots usign continuous shades.}
